Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-05-25 src/tex/ms.tex
update plots

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148_quadrant}). The
training waveforms used for fitting are listed in Table~\ref{tab:quadrants}. The
top-left and bottom-right regions have limited data for $q>4$, hence the results are
only valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that 
waveforms lying in the lower diagonal have mismatches worsened while other waveforms 
generally improve. For the top-right and bottom-left regions, we see that the optimization prefers waveforms with $\chi_1<\chi_2$. This further  In the top-left region, the improvement is significant, except for a
few defects caused by some testing waveforms having $q>4$. This can be seen clearer in 
Fig.~\ref{fig:all_quadrants}, which we see that the optimized histogram shifts downward 
rather uniformly. The optimized waveforms in the bottom-right region have two distinct 
features. Within this region, waveforms in the bottom-right corner improves significantly 
while others worsened. This phenomenon can also be seen in Fig.~\ref{fig:all_quadrants}, 
where the mismatch distribution shifted upward. Hence, we can deduce that the loss function 
descended into a different minimum compared to Fig.~\ref{fig:ps_q148}, making the optimized 
coefficients not useful in GW analysis. THus, further optimizing in a smaller region does 
not generally improve the results.
%\kw{I am a bit confused about why you can get worse result.}

@@

We have shown the result of recalibrating waveform coefficients. One thing to
note is that our recalibration procedure is not exactly the same as the original
calibration. For instance, we use a different set of NR waveforms, frequency
range, etc. Nonetheless, as the decrease in mismatch is rather significant, this
optimization procedure should be able to improve the accuracy of IMRPhenomD on a
similar scale regardless of the differences. Here, the result serves as a
demonstration of the general method used.  
% In Fig.~\ref{fig:0154}, the error in the inspiral region is around halved for
% both amplitude and phase. Similarly, for the merger-ringdown region, the
% optimized waveforms also show great improvement.

@@

The reduced spin approximation is a major contributor to the inaccuracies
observed in the ansatz. In IMRPhenomD, this approximation employs a single spin
parameter, $\chi_{\mathrm{PN}}$, as described in Sec. \ref{sec:method}. The
parameterization of BBH mergers using a single spin parameter results in a
degeneracy within the parameter space. Specifically, black hole events with
different spins may generate the same waveform due to identical values of
$\chi_{\mathrm{PN}}$, leading to erroneous results, particularly for highly
unequal spin events. This degeneracy produces straight lines in the parameter
space with negative slopes that depend on the mass ratio, as shown in
Fig.\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
region than the bottom-right region. In an attempt to address
this issue, we partitioned the parameter space into four regions, as described
in Sec. \ref{sec:result}. With separate optimizations for each
regions, Fig.\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
performance has improved, while the bottom-right region exhibits two distinct 
features. This is because the loss function has descended into a different 
minimum which highly favors events with $\chi_1\approx1$ and $\chi_2\approx-1$. 
This observation suggests the ansatz is specific to certain regions of the 
parameter space, with a preference for BBH events lying in the upper diagonal 
region, and it hardly fits with waveforms lying in the lower diagonal region.

The division of the parameter space into four regions was a simple approach
taken for practical reasons. A more systematic approach would involve the use of
level set estimation algorithms to identify regions of interest within the
parameter space. Such an algorithm can reveal additional degeneracies or issues
that may exist within the ansatz. One possible strategy is to recalibrate
individual regions of interest to achieve better results. An alternative
approach is to select regions based on degeneracy lines. However, due to the
limited number of NR waveforms available, we were unable to implement this
approach. With more NR waveforms available in the future that cover the entire
parameter space, we can perform optimization with fewer restrictions and select
regions more systematically.

@@

Furthermore, we investigated how the intrinsic parameters affect the
improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
certain preference for waveforms lying in the top-left region while the bottom-right 
region hardly improved. To further test this result, we recalibrate
waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
we can see that this recalibration process gives further
improvement to the top-left region while the bottom-right region falls into a different 
loss minimum. This indicates that the ansatz does not fit certain waveforms in this 
region, hence introduces bias to downstream GW analysis. This phenomenon is mainly 
due to the reduced spin approximation used in parameterizing the 
ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.

@@

% From analyzing mismatches in parameter space, we investigated the accuracy of
% generated waveforms based on their mass ratio and spin. Similar studies can be
% performed on other GW models to give better criteria to

% Improved the IMRPhenomD model. Found regions that can be improved. Checked
% ansatz.

>>>>

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148_quadrant}). 
The training waveforms used for fitting are listed in Tab.~\ref{tab:quadrants}. The
top-left and bottom-right regions have limited data for $q>4$, hence the results are
only valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that 
equal-spin waveforms lying on the diagonal has great improvement. Above the diagonal, 
the improvement is significant, except for a few defects caused by some testing waveforms 
having $q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see 
that the optimized histogram shifts downward uniformly. Below the diagonal, waveforms 
have negligible improvement (Fig.~\ref{fig:all_quadrants}~and~\ref{fig:ps_q148_quadrant}), 
indicating the original set of coefficients is the optimal set of coefficients, and 
cannot be further improved. Hence, we can deduce IMRPhenomD has better match with the 
top-left region and does not fit well with waveforms lying in the bottom-right 
region. Thus, the ansatz restricts additional improvements and further optimizing in a 
smaller region does not generally improve the results.

@@

We have shown the result of recalibrating waveform coefficients. One thing to
note is that our recalibration procedure is not exactly the same as the original
calibration. For instance, we use a different set of NR waveforms, frequency
range, etc. Nonetheless, as the decrease in mismatch is rather significant, this
optimization procedure should be able to improve the accuracy of IMRPhenomD on a
similar scale regardless of the differences. Here, the result serves as a
demonstration of the general method used.

@@

The reduced spin approximation is a major contributor to the inaccuracies
observed in the ansatz. In IMRPhenomD, this approximation employs a single spin
parameter, $\chi_{\mathrm{PN}}$, as described in Sec. \ref{sec:method}. The
parameterization of BBH mergers using a single spin parameter results in a
degeneracy within the parameter space. Specifically, black hole events with
different spins may generate the same waveform due to identical values of
$\chi_{\mathrm{PN}}$, leading to erroneous results, particularly for highly
unequal spin events. This degeneracy produces straight lines in the parameter
space with negative slopes that depend on the mass ratio, which can be seen in 
Fig.\ref{fig:ps_q148}. Notably, the ansatz performs better in the top-left
region than the bottom-right region. In an attempt to address
this issue, we partitioned the parameter space into four regions, as described
in Sec. \ref{sec:result}. With separate optimizations for each
regions, Fig.\ref{fig:ps_q148_quadrant} indicates that the top-left region's 
performance has improved, while the bottom-right region hardly improves. 
This observation suggests the ansatz is specific to certain regions of the 
parameter space, with a preference for BBH events lying above the diagonal, 
and it has limited enhancement for events lying below the diagonal.

The division of the parameter space into four regions was a simple approach
taken for practical reasons. A more systematic approach would involve the use of
level set estimation algorithms to identify regions of interest within the
parameter space. Such an algorithm can reveal additional degeneracies or issues
that may exist within the ansatz. One possible strategy is to recalibrate
individual regions of interest to achieve better results. An alternative
approach is to select regions based on degeneracy lines. However, due to the
limited number of NR waveforms available, we were unable to implement this
approach. With more NR waveforms available in the future that cover the entire
parameter space, we can perform optimization with fewer restrictions and select
regions more systematically. Other than how to divide regions of interest, the 
choice of training waveforms also affects the final results greatly. Note that 
the choice of training waveforms listed in Tab.~\ref{tab:quadrants} were taken 
arbitrarily to test the effectiveness of separate fitting. Hence, if one takes 
a different set of training waveforms over the parameter space, the result might 
give additional features that can test and explain IMRPhenomD better.

@@

Furthermore, we investigated how the intrinsic parameters affect the
improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
certain preference for waveforms lying in the top-left region while the bottom-right 
region hardly improved. To further test this result, we recalibrate
waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
we can see that this recalibration process gives further
improvement to the top-left region while the bottom-right region only have little 
improvement. This indicates that the ansatz has limited match in this region and 
does not fit most waveforms in this region, hence introduces bias to downstream GW 
analysis. This phenomenon is mainly due to the reduced spin approximation used in 
parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.

@@
